---
title: "J2J Report"
author: "Lars Vilhuber"
date: "March 8, 2017"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## A toy news release

This is a draft news release for a Census Bureau product called the Job-to-Job (J2J) Flows. It plays around with dynamic text and just basic replicability.

### Setting parameters
First, we need to set some parameters and get the data (if this were a Shiny webapp, that might be a pulldown menu, in particular for the state). 

```{r define_parms}
j2j.vintage <- "R2016Q3"
# j2jtype <- "j2j"
j2j.type <- "j2jr"
j2j.seas <- "s"
j2j.state <- "us"
# the data area
j2j.urlbase <- "https://lehd.ces.census.gov/data/j2j"
j2j.url <- paste(j2j.urlbase,j2j.vintage,j2j.type,j2j.state,sep = "/")
j2j.file <- paste(j2j.type,j2j.state,"all.csv.gz",sep="_")
# select all sorts of demo and firm vars
# no testing if they are valid!
j2j.industry <- "00" 
j2j.sex <- "0"
j2j.agegrp <- "A00"
j2j.race <- "A0"
j2j.ethnicity <- "A0"
j2j.education <- "E0"
j2j.firmage <- "0"
j2j.firmsize <- "0"
# or we could simply select the agg_level
# this is the same as all of the above
j2j.agg_level <- "1"
j2j.plotvar <- "EEHireR"
# and source statement
j2j.source <- paste("U.S. Census Bureau, Job-to-Job Flows, Release",j2j.vintage,sep = " ")
# our preferred format for pct numbers
j2j.fmtpct <- "%1.1f"
```

We are going to work with seasonally `r ifelse(j2j.seas == "s",print(""),print("un"))`adjusted `r ifelse(j2j.type == "j2jr",print("rates"),print("levels"))` from the `r j2j.vintage` release.

### Getting the data
Getting the data implies downloading the file "`r j2j.file`" from `r j2j.url`.
```{r readj2j}
# we need the full data
conr <- gzcon(url(paste(j2j.url,j2j.file,sep="/")))
txt <- readLines(conr)
j2jdata <- read.csv(textConnection(txt))
j2jdata$yqstr <- paste(j2jdata$year,j2jdata$quarter,sep="Q")
j2jdata$linyq <- j2jdata$year + (j2jdata$quarter-1)/4
# we also need some metadata. for later.
```

## Doing stuff
We now have the data and can do stuff with it.

### The graph
Now we simply plot this:

#### Code for the graph
```{r prep_figure1, echo=FALSE}
library(ggplot2)
library(ggthemes)
library(stringr)

analysis <- subset(j2jdata,year > 2001 & str_to_lower(seasonadj) == j2j.seas & agg_level == j2j.agg_level)
plotdata <- analysis[,c("year","quarter",j2j.plotvar,"seasonadj","agg_level")]
# scale up the fractions
#plotdata[3] <- plotdata[3]*100
plotdata$yq <- plotdata$year + (plotdata$quarter-1)/4
gg <- ggplot(plotdata,aes_q(x=as.name("yq"),y=as.name(j2j.plotvar))) +
  geom_line() + 
  geom_point() +
  scale_x_continuous(breaks=seq(min(plotdata$year), max(plotdata$year), 1)) +
  scale_y_continuous(labels = scales::percent,
                     breaks=seq(0, max(plotdata[,j2j.plotvar]+0.005), 0.005),
                     limits=c(0,max(plotdata[,j2j.plotvar]+0.005))
                     ) +
  theme_economist_white() +
  theme(
    axis.title.y=element_blank(),
    axis.title=element_text(size=8,face="italic",hjust=1),
    axis.text.x=element_text(angle=+45,hjust=-0.02),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle =element_text(hjust = 0.5),
  ) +
  xlab(paste("Source:",j2j.source,sep = " ")) +
  ggtitle("Worker separations to new jobs within the quarter", "seasonally adjusted") 
```
#### The actual graph
```{r figure1, echo=FALSE}
gg
```

### The text
Now we can reference some of the data as well.

#### Preparing the text
```{r prep_analysis}
j2j.lyq <- max(analysis$yqstr)
j2j.ly <- subset(analysis,yqstr == j2j.lyq)[1,"year"]
j2j.lq <- subset(analysis,yqstr == j2j.lyq)[1,"quarter"]
j2j.linyq <- j2j.ly + (j2j.lq -1)/4

j2j.refword <- function(cardinal) {
  if ( cardinal == 1 ) tmpsuffix <- "first"
  if ( cardinal == 2 ) tmpsuffix <- "second"
  if ( cardinal == 3 ) tmpsuffix <- "third"
  if ( cardinal == 4 ) tmpsuffix <- "fourth"
  return(tmpsuffix)
}

j2j.compare <- function(this,last) {
  comp <- (this - last)/last * 100
  if ( comp > 0 ) word <- "grew"
  if ( comp < 0 ) word <- "shrunk"
  # 1 percent difference in numbers is unchanged
  if ( abs(comp) < 1 ) word <- "were largely unchanged"
  return(word)
}
# pick our poison
j2j.current <- subset(analysis,linyq==j2j.linyq)
j2j.prevq <- subset(analysis,linyq==j2j.linyq - 0.25)
j2j.prevy <- subset(analysis,linyq==j2j.linyq - 1)

```

## The actual text
The fraction of workers changing jobs 
`r j2j.compare(j2j.current[,j2j.plotvar],j2j.prevy[,j2j.plotvar])`
in the `r j2j.refword(j2j.lq)` quarter of `r j2j.ly`,
with `r sprintf(j2j.fmtpct,round(j2j.current[,j2j.plotvar]*100,1))`% changing employers this quarter, 
compared to 
`r sprintf(j2j.fmtpct,round(j2j.prevy[,j2j.plotvar]*100,1))`% 
in the `r j2j.refword(j2j.lq)` quarter of `r j2j.ly - 1`.
Flows into employment
`r j2j.compare(j2j.current[,"NEHireR"],j2j.prevy[,"NEHireR"])`, 
with `r sprintf(j2j.fmtpct,round(j2j.current[,"NEHireR"]*100,1))`% 
employed on the last day of the quarter who did not hold a job on the first day of the quarter,
compared to `r sprintf(j2j.fmtpct,round(j2j.prevy[,"NEHireR"]*100,1))`% a year ago.
Separations to non-employment 
`r j2j.compare(j2j.current[,"ENSepR"],j2j.prevy[,"ENSepR"])`, 
with `r sprintf(j2j.fmtpct,round(j2j.current[,"ENSepR"]*100,1))`% 
employed on the first day of the quarter and no longer employed on the last day of the quarter,
compared to `r sprintf(j2j.fmtpct,round(j2j.prevy[,"ENSepR"]*100,1))`% a year ago.


